19Q1

### 2019-05-08 19Q1.Rmd
updated Rmd

Mutation calls (coding region, germline filtered):
```{r mutation_status, eval=FALSE}
maf_raw <- read.delim("./data_munging/19Q1/depmap_19Q1_mutation_calls_20190305.csv.gz", header = TRUE, sep = ",", row.names = 1)
# Harmonize ID column name
colnames(maf_raw)[colnames(maf_raw) == "DepMap_ID"] <- "Broad_ID"
colnames(maf_raw)[colnames(maf_raw) == "Hugo_Symbol"] <- "Gene"
# Select columns
maf_df <- subset(maf_raw, select = c("Gene", "Broad_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1"))
maf_df$Reference_Allele <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1 <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
maf_df <- maf_df[maf_df$Variant_Classification != "Silent",]
# Add Mutation_Status column
maf_df$Mutation_Status <- 1
maf_df$PointMutation <- with(maf_df, ifelse(Reference_Allele == 1 | is.na(Reference_Allele), TRUE, ifelse(Reference_Allele == 0 & Tumor_Seq_Allele1 == 1, 1, 0)))

@@

DepMap WES CN Data:
```{r copy_number, eval=FALSE}
cn <- read.delim("./data_munging/19Q1/public_19Q1_gene_cn_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
# Remove Entrez gene IDs from colnames, format is "Gene (Entrez ID)"
colnames(cn) <- gsub(" .*", "", colnames(cn))
colnames(cn)[1] <- "Broad_ID"
# Melt
cn_melt <- melt(data = cn, id.vars = "Broad_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Gene", value.name = "CN")

@@

19Q1, CCLE RNAseq gene expression data (log2(TPM+1)):
```{r gene_expression, eval=FALSE}
ge <- read.delim("./../htscreens_giant_files/CCLE_depMap_19Q1_TPM_20190305.csv.gz", header = TRUE, sep = ",", check.names = FALSE)
# Remove Entrez gene IDs from colnames, format is "Gene (Entrez ID)"
colnames(ge) <- gsub(" .*", "", colnames(ge_19Q1))
colnames(ge)[1] <- "Broad_ID"
# Melt
ge_melt <- melt(ge, id.vars = "Broad_ID", measure.vars = colnames(ge_19Q1)[2:ncol(ge_19Q1)], variable.name = "Gene", value.name = "TPM")

@@

    # Get all G2P gene-cell line combinations
    grid <- expand.grid("Cell_line" = unique(filt$Cell_line), "Gene" = g2p_genes)
    data <-  merge(filt, grid, by = "Cell_line", all.y = TRUE)
    print("Merged grid.")
    # Merge metadata
    data <-  merge(data, select(ccl_meta, Cell_line, Histology, CCLE_Name, Broad_ID, Primary.Disease), by = c("Cell_line", "Histology", "CCLE_Name", "Broad_ID"), all.x = TRUE)
    ## Add columns for G2P drug-gene and drug-gene-lineage sorting
    data$Drug_Gene <- paste0(data$Drug, "_", data$Gene)
    data$G2P_Drug_Gene <- ifelse(data$Drug_Gene %in% g2p_druggenes, "Yes", "No")
    data$Drug_Gene_Lin <- paste0(data$Drug, "_", data$Gene, "_", data$Primary.Disease)
    data$G2P_Drug_Gene_Lin <- ifelse(data$Drug_Gene_Lin %in% g2p_dgls, "Yes", "No")
    print("Merged metadata.")
  }
  if(database == "CRISPR") {
    print(paste0("Processing ", database, " data."))
    # Melt CRISPR dataset for merging
    data <- melt(crispr_all, id.vars = "Broad_ID", measure.vars = colnames(crispr_all)[2:ncol(crispr_all)], variable.name = "Gene", value.name = "Score")
    # Get all G2P gene-cell line combinations
    data <-  filter(data, Gene %in% g2p_genes)
    # Merge cell line metadata
    data <- merge(data, select(ccl_meta, Cell_line, Histology, CCLE_Name, Broad_ID, Primary.Disease), by = "Broad_ID", all.x = TRUE)
    data$Drug_Gene <- paste0(data$Drug, "_", data$Gene)
    data$Drug_Gene_Lin <- paste0(data$Drug, "_", data$Gene, "_", data$Primary.Disease)
    print("Merged metadata.")
  }

  # Merge mutation calls
  data <- merge(data, maf_df_filt, by = c("Gene", "Broad_ID"), all.x = TRUE)
  data$Mutation_Status <- ifelse(is.na(data$Mutation_Status), 0, data$Mutation_Status)
  ## Identify point mutations
  data$PointMutation <- ifelse(is.na(data$PointMutation), 1, data$PointMutation)
  ## Summarize number of mutant and Wildtype cell lines
  data_muts_summ <- data %>% group_by(Gene) %>% summarize(N_Wildtype = length(Mutation_Status) - sum(Mutation_Status == 1), N_Mutant = sum(Mutation_Status == 1))
  ## Merge test results back into full dataset, which restores information lost in the summarization
  data <- merge(data_muts_summ, data, by = "Gene")
  ## Add color columns for plotting
  if(database == "CCLE") {
    data$Color <- ifelse(data$Mutation_Status == 0, "palegreen3", "springgreen4")
  }
  if(database == "CTRP") {
    data$Color <- ifelse(data$Mutation_Status == 0, "slategray3", "steelblue4")
  }
  if(database == "GDSC") {
    data$Color <- ifelse(data$Mutation_Status == 0, "paleturquoise3", "turquoise4")
  }
  if(database == "CRISPR") {
    data$Color <- ifelse(data$Mutation_Status == 0, "thistle", "darkorchid")
  }
  data$Color <- factor(data$Color)
  print("Merged mutation calls.")

@@

  # Drop cell lines not in DepMap
  data <- filter(data, !is.na(Broad_ID))

@@

FilterForPointMuts <- function(dataframe, outfile) {
  # Filter for point mutations
  ptmuts <- filter(dataframe, PointMutation == 1)
  # Remove columns that lead to duplicate mutation entries
  ptmuts <- subset(ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, PointMutation))
  ptmuts$Gene <- ordered(ptmuts$Gene, levels = g2p_genes)
  ptmuts <- unique(ptmuts)
  saveRDS(ptmuts, outfile, compress = "xz")
  return(ptmuts)
}
```

@@

`r nrow(filter(test, n != 1))`/`r nrow(test)` gene-cell line combinations have more than one mutation in the MAF file. This is corrected in the point mutation filtering done below.
```{r merge_filter_drug_data, eval=FALSE}
# Load in Pozdeyev et al. (2016) data
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")

@@

ccle <- MergeOmicsG2P(database = "CCLE", outfile = "./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
ctrp <- MergeOmicsG2P(database = "CTRP", outfile = "./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes.rds")
gdsc <- MergeOmicsG2P(database = "GDSC", outfile = "./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")

ccle_ptmuts <- FilterForPointMuts(dataframe = ccle, outfile = "./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
ctrp_ptmuts <- FilterForPointMuts(dataframe = ctrp, outfile = "./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- FilterForPointMuts(dataframe = gdsc, outfile = "./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")
```

```{r load_drug_rds, eval=TRUE}
ccle <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
ctrp <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes.rds")
gdsc <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")
ccle_ptmuts <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
ctrp_ptmuts <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")

@@

Merge all CRISPR genes:
```{r merge_crispr_data, eval=FALSE}
crispr_all <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr_all) <- gsub(" .*", "", colnames(crispr_all))
colnames(crispr_all)[1] <- "Broad_ID"

@@

# Merge mutation annotations
maf_filt_crispr <- filter(maf_df, Gene %in% as.character(unique(crispr_melt$Gene)))
maf_filt_crispr <- subset(maf_filt_crispr, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1))
maf_filt_crispr <- unique(maf_filt_crispr)
crispr_muts <- merge(crispr_melt, maf_filt_crispr, by = c("Gene","Broad_ID"), all.x = TRUE)
crispr_muts$Gene <- factor(crispr_muts$Gene)
crispr_muts$Mutation_Status <- ifelse(is.na(crispr_muts$Mutation_Status), 0, crispr_muts$Mutation_Status)
## Identify point mutations and wildtypes for testing
crispr_muts$PointMutation <- ifelse(is.na(crispr_muts$PointMutation), 1, crispr_muts$PointMutation)

@@

saveRDS(crispr_data, "./data_munging/rds/crispr_data_19Q1.rds", compress = "xz")

# Filter G2P CRISPR for point mutations
crispr_ptmuts <- filter(crispr_data, PointMutation == 1)
crispr_ptmuts <- subset(crispr_ptmuts, select = -c(PointMutation))
crispr_ptmuts <- filter(crispr_ptmuts, Gene %in% g2p_genes)
crispr_ptmuts$Primary.Disease <- factor(crispr_ptmuts$Primary.Disease)
crispr_ptmuts$Mutation_Status <- factor(crispr_ptmuts$Mutation_Status)
saveRDS(crispr_ptmuts, "./data_munging/rds/crispr_data_19Q1_g2pgenes_ptmuts.rds", compress = "xz")
```

```{r load_crispr_filter, eval=TRUE}
crispr_data <- readRDS("./data_munging/rds/crispr_data_19Q1.rds")
crispr_ptmuts <- readRDS("./data_munging/rds/crispr_data_19Q1_g2pgenes_ptmuts.rds")
```

@@

@@

```{r ccl_venn, eval=FALSE}
ccle_ccls <- unique(ccle_ptmuts$Broad_ID)
ctrp_ccls <- unique(ctrp_ptmuts$Broad_ID)
gdsc_ccls <- unique(gdsc_ptmuts$Broad_ID)
crispr_ccls <- unique(crispr_ptmuts$Broad_ID)

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn.diagram(x = list(ccle_ccls, ctrp_ccls, gdsc_ccls, crispr_ccls),
             filename = "./plots/manuscript/venn_ccls.png",
             output = TRUE, imagetype = "png",
             height = 2000, width = 2000, resolution = 500,
             lwd = 1, cex = 1,
             category.names = c("CCLE", "CTRP", "GDSC", "CRISPR"),
             cat.cex = 1.4, cat.fontface = "bold", cat.default.pos = "outer",
             # cat.dist = c(0.06, 0.06, 0.035),
             fill = c("springgreen4", "steelblue4", "turquoise4", "darkorchid"))
```

```{r drug_venn, eval=FALSE}
ccle_drugs <- unique(ccle_ptmuts$Drug)
ctrp_drugs <- unique(ctrp_ptmuts$Drug)
gdsc_drugs <- unique(gdsc_ptmuts$Drug)

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn.diagram(x = list(ccle_drugs, ctrp_drugs, gdsc_drugs),
             filename = "./plots/manuscript/venn_drugs.png",
             output = TRUE, imagetype = "png",
             height = 2000, width = 2000, resolution = 500,
             lwd = 1, cex = 1.4,
             category.names = c("CCLE", "CTRP", "GDSC"),
             cat.cex = 1.4, cat.fontface = "bold", cat.default.pos = "outer",
             cat.dist = c(0.06, 0.06, 0.035),
             fill = c("palegreen3", "slategray3", "paleturquoise3"))
```

@@

```{r histology_dists_data, eval=FALSE}
ccle_hists <- unique(select(ccle_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
ccle_hists$Perc <- ccle_hists$n / sum(ccle_hists$n) * 100
ccle_hists$Dataset <- "CCLE"
ctrp_hists <- unique(select(ctrp_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
ctrp_hists$Perc <- ctrp_hists$n / sum(ctrp_hists$n) * 100
ctrp_hists$Dataset <- "CTRP"
gdsc_hists <- unique(select(gdsc_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
gdsc_hists$Perc <- gdsc_hists$n / sum(gdsc_hists$n) * 100
gdsc_hists$Dataset <- "GDSC"
crispr_hists <- unique(select(crispr_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
crispr_hists$Perc <- crispr_hists$n / sum(crispr_hists$n) * 100
crispr_hists$Dataset <- "CRISPR"

drug_hists <- rbind(ccle_hists, ctrp_hists, gdsc_hists, crispr_hists) %>% drop_na(Primary.Disease)
drug_hists$Dataset <- factor(drug_hists$Dataset, levels = c("CCLE", "CTRP", "GDSC", "CRISPR"))
drug_hists$Primary.Disease <- as.character(drug_hists$Primary.Disease)
drug_hists$Primary.Disease <- gsub(" cancer", "", drug_hists$Primary.Disease)
# saveRDS(drug_hists, "./data_munging/rds/histology_distributions.rds")
```

```{r histology_dists, fig.height=4.5, fig.width=4}
drug_hists <- readRDS("./data_munging/rds/histology_distributions.rds")
png(filename = "./plots/manuscript/histology_distributions.png", width = 8, height = 9, units = "in", res = 500)
ggplot(drug_hists, aes(x = Dataset, y = n, fill = Primary.Disease)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(breaks = seq(0, 850, by = 50), labels = seq(0, 850, by = 50)) +
  guides(fill = guide_legend(title = "Histology", ncol = 1)) +
  labs(y = "Count")
dev.off()
```

@@

# png(filename = "./plots/manuscript/g2p_summary_heatmap.png", width = 10, height = 9, units = "in", res = 500)
draw(hta + ht, gap = unit(5, "mm"), heatmap_legend_side = "top")
# dev.off()
```

@@

```{r wilcoxon_tests, eval=FALSE}
crispr_ptmuts <- unfactorize(crispr_ptmuts)
crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

@@

crispr_gene_mut_plot <- ggplot(data = crispr_ptmuts) +
  facet_wrap(~ Gene, nrow = 2, drop = FALSE) +
  scale_color_manual(name = "Mutation Status", labels = c("Wildtype", "Mutant"), values = c("1" = "darkorchid", "0" = "thistle")) +
  scale_x_discrete(name = "Mutation Status", labels = c("0" = "WT", "1" = "Mut")) +
  geom_jitter(aes(x = Mutation_Status, y = Score, color = Mutation_Status), width = 0.2, alpha = 0.5, size = 0.6) +
  geom_boxplot(aes(x = Mutation_Status, y = Score), fill = "transparent", outlier.shape = NA, lwd = 0.3) +
  geom_text(data = crispr_signif_gene, aes(x = 1.5, y = 1.75, label = p.signif.adj)) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  theme(legend.position = "top", legend.direction = "horizontal") +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  labs(y = "CERES Score")
# crispr_gene_mut_plot

# ggsave("./plots/manuscript/crispr_gene_mut_plot.png", crispr_gene_mut_plot, width = 12, height = 6, units = "in")
```

>>>>

Mutation calls (coding region, germline filtered):
```{r mutation_status, eval=FALSE}
maf_raw <- read.delim("./data_munging/19Q1/depmap_19Q1_mutation_calls_20190305.csv.gz", header = TRUE, sep = ",", row.names = 1)
# Harmonize ID column name
colnames(maf_raw)[colnames(maf_raw) == "DepMap_ID"] <- "Broad_ID"
colnames(maf_raw)[colnames(maf_raw) == "Hugo_Symbol"] <- "Gene"
# Select columns
maf_df <- subset(maf_raw, select = c("Gene", "Broad_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1"))
maf_df$Reference_Allele <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1 <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
maf_df <- maf_df[maf_df$Variant_Classification != "Silent",]
# Add Mutation_Status column
maf_df$Mutation_Status <- 1
maf_df$PointMutation <- with(maf_df, ifelse(Reference_Allele == 1 | is.na(Reference_Allele), TRUE, ifelse(Reference_Allele == 0 & Tumor_Seq_Allele1 == 1, 1, 0)))
maf_df <- unfactorize(maf_df)

@@

DepMap WES CN Data:
```{r copy_number, eval=FALSE}
cn <- read.delim("./data_munging/19Q1/public_19Q1_gene_cn_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
# Remove Entrez gene IDs from colnames, format is "Gene (Entrez ID)"
colnames(cn) <- gsub(" .*", "", colnames(cn))
colnames(cn)[1] <- "Broad_ID"
# Melt
cn_melt <- melt(data = cn, id.vars = "Broad_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Gene", value.name = "CN")
cn_melt <- unfactorize(cn_melt)

@@

19Q1, CCLE RNAseq gene expression data (log2(TPM+1)):
```{r gene_expression, eval=FALSE}
ge <- read.delim("./../htscreens_giant_files/CCLE_depMap_19Q1_TPM_20190305.csv.gz", header = TRUE, sep = ",", check.names = FALSE)
# Remove Entrez gene IDs from colnames, format is "Gene (Entrez ID)"
colnames(ge) <- gsub(" .*", "", colnames(ge))
colnames(ge)[1] <- "Broad_ID"
# Melt
ge_melt <- melt(ge, id.vars = "Broad_ID", measure.vars = colnames(ge)[2:ncol(ge)], variable.name = "Gene", value.name = "TPM")
ge_melt$TPM <- log2(ge_melt$TPM + 1)
ge_melt <- unfactorize(ge_melt)

@@

    # Get all G2P gene-cell line combinations
    grid <- expand.grid("Cell_line" = unique(filt$Cell_line), "Gene" = g2p_genes)
    data <-  merge(filt, grid, by = "Cell_line", all.y = TRUE)
    print("Merged grid.")
    # Merge metadata
    data <-  merge(data, select(ccl_meta, Cell_line, Histology, CCLE_Name, Broad_ID, Primary.Disease), by = c("Cell_line", "Histology", "CCLE_Name", "Broad_ID"), all.y = TRUE)
    ## Add columns for G2P drug-gene and drug-gene-lineage sorting
    data$Drug_Gene <- paste0(data$Drug, "_", data$Gene)
    data$G2P_Drug_Gene <- ifelse(data$Drug_Gene %in% g2p_druggenes, "Yes", "No")
    data$Drug_Gene_Lin <- paste0(data$Drug, "_", data$Gene, "_", data$Primary.Disease)
    data$G2P_Drug_Gene_Lin <- ifelse(data$Drug_Gene_Lin %in% g2p_dgls, "Yes", "No")
    print("Merged metadata.")
  }
  if(database == "CRISPR") {
    print(paste0("Processing ", database, " data."))
    # Melt CRISPR dataset for merging
    data <- melt(crispr_all, id.vars = "Broad_ID", measure.vars = colnames(crispr_all)[2:ncol(crispr_all)], variable.name = "Gene", value.name = "Score")
    # Get all G2P gene-cell line combinations
    data <-  filter(data, Gene %in% g2p_genes)
    # Merge cell line metadata
    data <- merge(data, select(ccl_meta, Cell_line, Histology, CCLE_Name, Broad_ID, Primary.Disease), by = "Broad_ID", all.y = TRUE)
    data$Drug_Gene <- paste0(data$Drug, "_", data$Gene)
    data$Drug_Gene_Lin <- paste0(data$Drug, "_", data$Gene, "_", data$Primary.Disease)
    print("Merged metadata.")
  }

  # Merge mutation calls
  data <- unfactorize(data)
  data <- merge(data, maf_df_ptmuts, by = c("Gene", "Broad_ID"), all.x = TRUE)
  data$Mutation_Status <- ifelse(is.na(data$Mutation_Status), 0, data$Mutation_Status)
  ## Identify point mutations
  # data$PointMutation <- ifelse(is.na(data$PointMutation), 1, data$PointMutation)
  ## Summarize number of mutant and Wildtype cell lines
  data_muts_summ <- data %>% group_by(Gene) %>% summarize(N_Wildtype = length(Mutation_Status) - sum(Mutation_Status == 1), N_Mutant = sum(Mutation_Status == 1))
  ## Merge test results back into full dataset, which restores information lost in the summarization
  data <- merge(data_muts_summ, data, by = "Gene")
  print("Merged mutation calls.")

@@

  # Drop cell lines not in DepMap
  data <- filter(data, !is.na(Broad_ID) & !(Broad_ID %in% c("ACH-001182", "ACH-001451")))

@@

# FilterForPointMuts <- function(dataframe, outfile) {
#   # Filter for point mutations
#   ptmuts <- filter(dataframe, PointMutation == 1)
#   # Remove columns that lead to duplicate mutation entries
#   ptmuts <- subset(ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, PointMutation))
#   ptmuts$Gene <- ordered(ptmuts$Gene, levels = g2p_genes)
#   ptmuts <- unique(ptmuts)
#   saveRDS(ptmuts, outfile, compress = "xz")
#   return(ptmuts)
# }
```

@@

`r nrow(filter(test, n != 1))`/`r nrow(test)` gene-cell line combinations have more than one mutation in the MAF file. This is corrected in the point mutation filtering done below.

**RE-FILTER FOR PTMUTS**
```{r merge_filter_drug_data, eval=FALSE}
# Load in Pozdeyev et al. (2016) data
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")

@@

# Omics metadata
maf_df_ptmuts <- unique(filter(maf_df, Gene %in% g2p_genes & PointMutation == 1) %>% select(-c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, PointMutation)))
ge_filt <- filter(ge_melt, Gene %in% g2p_genes)
cn_filt <- filter(cn_melt, Gene %in% g2p_genes)

ccle_ptmuts <- MergeOmicsG2P(database = "CCLE", outfile = "./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
ctrp_ptmuts <- MergeOmicsG2P(database = "CTRP", outfile = "./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- MergeOmicsG2P(database = "GDSC", outfile = "./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")
```

```{r load_drug_rds, eval=TRUE}
ccle_ptmuts <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
ctrp_ptmuts <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")

@@

Merge all CRISPR genes:
```{r merge_crispr_data, eval=TRUE}
crispr_all <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr_all) <- gsub(" .*", "", colnames(crispr_all))
colnames(crispr_all)[1] <- "Broad_ID"

@@

# Merge mutation annotations
maf_filt_crispr <- unique(filter(maf_df, Gene %in% as.character(unique(crispr_melt$Gene)) & PointMutation == 1) %>% select(-c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, PointMutation)))
crispr_muts <- merge(crispr_melt, maf_filt_crispr, by = c("Gene","Broad_ID"), all.x = TRUE)
crispr_muts$Mutation_Status <- ifelse(is.na(crispr_muts$Mutation_Status), 0, crispr_muts$Mutation_Status)

@@

saveRDS(crispr_data, "./data_munging/rds/crispr_data_19Q1_ptmuts.rds", compress = "xz")

# Filter G2P CRISPR for point mutations
# crispr_ptmuts <- filter(crispr_data, PointMutation == 1)
# crispr_ptmuts <- subset(crispr_ptmuts, select = -c(PointMutation))
# crispr_ptmuts <- filter(crispr_ptmuts, Gene %in% g2p_genes)
# crispr_ptmuts$Primary.Disease <- factor(crispr_ptmuts$Primary.Disease)
# crispr_ptmuts$Mutation_Status <- factor(crispr_ptmuts$Mutation_Status)
# saveRDS(crispr_ptmuts, "./data_munging/rds/crispr_data_19Q1_g2pgenes_ptmuts.rds", compress = "xz")
```

```{r load_crispr_filter, eval=TRUE}
# crispr_data <- readRDS("./data_munging/rds/crispr_data_19Q1.rds")
crispr_ptmuts <- readRDS("./data_munging/rds/crispr_data_19Q1_ptmuts.rds")
```

@@

## Mutation

@@

```{r ccl_venn, eval=TRUE}
ccle_ccls <- unique(ccle_ptmuts$Broad_ID)
ctrp_ccls <- unique(ctrp_ptmuts$Broad_ID)
gdsc_ccls <- unique(gdsc_ptmuts$Broad_ID)
crispr_ccls <- unique(crispr_ptmuts$Broad_ID)

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn.diagram(x = list(ccle_ccls, ctrp_ccls, gdsc_ccls, crispr_ccls),
             filename = "./plots/manuscript/all_venn_ccls.png",
             output = TRUE, imagetype = "png",
             height = 2000, width = 2000, resolution = 500,
             lwd = 1, cex = 1,
             category.names = c("CCLE", "CTRP", "GDSC", "CRISPR"),
             cat.cex = 1.4, cat.fontface = "bold", cat.default.pos = "outer",
             # cat.dist = c(0.06, 0.06, 0.035),
             fill = c("springgreen4", "steelblue4", "turquoise4", "darkorchid"))

venn.diagram(x = list(ccle_ccls, ctrp_ccls, gdsc_ccls),
             filename = "./plots/manuscript/drug_venn_ccls.png",
             output = TRUE, imagetype = "png",
             height = 2000, width = 2000, resolution = 500,
             lwd = 1, cex = 1.4,
             category.names = c("CCLE", "CTRP", "GDSC"),
             cat.cex = 1.4, cat.fontface = "bold", cat.default.pos = "outer",
             cat.dist = c(0.06, 0.06, 0.035),
             fill = c("springgreen4", "steelblue4", "turquoise4"))
```

```{r drug_venn, eval=TRUE}
ccle_drugs <- unique(ccle_ptmuts$Drug)
ctrp_drugs <- unique(ctrp_ptmuts$Drug)
gdsc_drugs <- unique(gdsc_ptmuts$Drug)

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn.diagram(x = list(ccle_drugs, ctrp_drugs, gdsc_drugs),
             filename = "./plots/manuscript/drug_venn_drugs.png",
             output = TRUE, imagetype = "png",
             height = 2000, width = 2000, resolution = 500,
             lwd = 1, cex = 1.4,
             category.names = c("CCLE", "CTRP", "GDSC"),
             cat.cex = 1.4, cat.fontface = "bold", cat.default.pos = "outer",
             cat.dist = c(0.06, 0.06, 0.035),
             fill = c("palegreen3", "slategray3", "paleturquoise3"))
```

@@

```{r histology_dists_data, eval=TRUE}
ccle_hists <- unique(select(ccle_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
ccle_hists$Perc <- ccle_hists$n / sum(ccle_hists$n) * 100
ccle_hists$Dataset <- "CCLE"
ctrp_hists <- unique(select(ctrp_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
ctrp_hists$Perc <- ctrp_hists$n / sum(ctrp_hists$n) * 100
ctrp_hists$Dataset <- "CTRP"
gdsc_hists <- unique(select(gdsc_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
gdsc_hists$Perc <- gdsc_hists$n / sum(gdsc_hists$n) * 100
gdsc_hists$Dataset <- "GDSC"
crispr_hists <- unique(select(crispr_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
crispr_hists$Perc <- crispr_hists$n / sum(crispr_hists$n) * 100
crispr_hists$Dataset <- "CRISPR"

drug_hists <- rbind(ccle_hists, ctrp_hists, gdsc_hists, crispr_hists) %>% drop_na(Primary.Disease)
drug_hists$Dataset <- factor(drug_hists$Dataset, levels = c("CCLE", "CTRP", "GDSC", "CRISPR"))
drug_hists$Primary.Disease <- as.character(drug_hists$Primary.Disease)
drug_hists$Primary.Disease <- gsub(" cancer", "", drug_hists$Primary.Disease)
saveRDS(drug_hists, "./data_munging/rds/histology_distributions.rds")
```

```{r histology_dists, fig.height=4.5, fig.width=4}
drug_hists <- readRDS("./data_munging/rds/histology_distributions.rds")
png(filename = "./plots/manuscript/all_histology_distributions.png", width = 8, height = 9, units = "in", res = 500)
ggplot(drug_hists, aes(x = Dataset, y = n, fill = Primary.Disease)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(breaks = seq(0, 850, by = 50), labels = seq(0, 850, by = 50)) +
  guides(fill = guide_legend(title = "Histology", ncol = 1)) +
  labs(y = "Count")
dev.off()
```

```{r, eval=TRUE}
ccle_ccls <- unique(ccle_ptmuts$Broad_ID)
ctrp_ccls <- unique(ctrp_ptmuts$Broad_ID)
gdsc_ccls <- unique(gdsc_ptmuts$Broad_ID)
crispr_ccls <- unique(crispr_ptmuts$Broad_ID)

ccls <- data.frame(Broad_ID = union(ccle_ccls, union(ctrp_ccls, gdsc_ccls)))
all_ccls <- merge(ccls, ccl_meta, by = "Broad_ID")
all_ccls$Primary.Disease <- gsub(" cancer", "", all_ccls$Primary.Disease)

png(filename = "./plots/manuscript/drug_histology_distributions.png", width = 8, height = 4, units = "in", res = 500)
ggplot(all_ccls, aes(x = fct_infreq(factor(Primary.Disease)), fill = Primary.Disease)) +
  geom_bar(color = "black") +
  scale_y_continuous(breaks = seq(0, 220, by = 40), labels = seq(0, 220, by = 40)) +
  guides(fill = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  labs(y = "Number of Cell Lines", x = "Histology", caption = paste0(formatC(nrow(all_ccls), format = "d", big.mark = ","), " cell lines represented."))
dev.off()
```

@@

png(filename = "./plots/manuscript/g2p_summary_heatmap.png", width = 10, height = 9, units = "in", res = 500)
draw(hta + ht, gap = unit(5, "mm"), heatmap_legend_side = "top")
dev.off()
```

@@

```{r wilcoxon_tests, eval=TRUE}
crispr_ptmuts <- unfactorize(crispr_ptmuts)
crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

@@

crispr_gene_mut_plot <- ggplot(data = crispr_ptmuts) +
  facet_wrap(~ Gene, nrow = 2, drop = FALSE) +
  scale_color_manual(name = "Mutation Status", labels = c("Wildtype", "Mutant"), values = c("1" = "darkorchid", "0" = "thistle")) +
  scale_x_discrete(name = "Mutation Status", labels = c("0" = "WT", "1" = "Mut")) +
  geom_jitter(aes(x = Mutation_Status, y = Score, color = Mutation_Status), width = 0.2, alpha = 0.5, size = 0.6) +
  geom_boxplot(aes(x = Mutation_Status, y = Score), fill = "transparent", outlier.shape = NA, lwd = 0.3) +
  geom_text(data = crispr_signif_gene, aes(x = 1.5, y = 1.75, label = p.signif)) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  theme(legend.position = "top", legend.direction = "horizontal") +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  labs(y = "CERES Score")
crispr_gene_mut_plot

ggsave("./plots/manuscript/crispr_gene_mut_plot.png", crispr_gene_mut_plot, width = 12, height = 6, units = "in")
```
